---
title: "Exploratory Data Analysis"
author: "JVWCD Capstone Group"
date: "2/20/2022"
output: 
  html_document:
    highlight: breezedark
    number_sections: yes
    toc: yes
    fig_width: 15
    fig_height: 10
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Business Statement

There is a strong correlation between water usage and the weather in the water industry. Water usage increases
during warmer and drier periods, and when it’s colder and wetter, water usage is typically less. Jordan Valley
Water Conservancy District (JVWCD) has noticed this trend and is wanting to find a solution that will allow
them to be more proactive instead of reactive with these seasonal trends.

Historically, operations staff have worked in a reactive state chasing water demands of community members,
which is influenced by changing weather. Most of this variability happens in the summer, with residential
outdoor landscape watering being the leading consumer. Roughly 65% of the community's water use is in that
outdoor landscape space. This reactive chase ensures that the community always has water, which often
requires water managers, supervisors and operators to spend long hours (including overtime hours), watching
and reacting to changes in water demand. Managing the system's water demand can be like steering a large
ship – sudden and abrupt changes are not easily absorbed. Also, there is a cost factor to this. Well-informed
planning to operate the system with consistency and minor tweaks saves money. However, operational costs
can add up quickly if a well is started at an unplanned time. This event could cost the water system $5,000 to
$9,000 in electricity demand charges alone.

JVWCD would like to be more efficient with its resources and manpower. The goal of this capstone project is
to explore modeling water demand against weather data. The ideal outcome of this project would be a
predictive model that gives the operations team more confidence in making decisions related to water demand
so that they can move to a more proactive form of water deliveries.


# Analytic Objectives

System demand is calculated by adding all the individual sources and subtracting the reservoir trends
from the total. JVWCD has provided system data to use for modeling. This is time series data and could
require some cleaning before use in a model.

JVWCD is currently using the online version of Tableau. It would be ideal if we had a model that could
be used there so as to reduce the amount of training it would take to implement this model. JVWCD pays
for 30 licenses of Tableau Online.

One aspect of the analytics problem is selecting a weather service to provide weather data. The one that
stands out to us right now is openweathermap.org and we’d likely get the information we need from the
free version but are also looking at other options to compare against. The team will look at two additional
sources of weather data to benchmark against the desired version (openweathermap.org).


# Exploratory Questions

The questions below were crafted prior to our EDA. We used these questions to help guide our EDA and added to it during our EDA. 

Is there strong correlations between the different weather metrics and the different point type?
Do we see a seasonal trend with the data?
Can the weather help predict demand?
How does weather change season by season?
Is there anything useful with hourly data vs. daily weather data?
How does demand change from month to month?
Is there a correlation between demand and temperature, and how strong is that correlation?
What are the best predictor variables to help predict daily demand?
Seasonal variation?


# Read Data

## Packages

```{r}
list.of.packages <- c("tidyverse",
                      "lubridate",
                      "chron",
                      "knitr",
                      "corrplot",
                      "caret",
                      "scales",
                      "gridExtra",
                      "useful",
                      "ISOweek")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(tidyverse)
library(lubridate)
library(chron)
library(knitr)
library(corrplot)
library(caret)
library(scales)
library(gridExtra)
library(useful)
library(ISOweek)
```

## Read CSV files and pre process

```{r}
#read in data. Change names to be more readable and make hourly demand numeric
sd <- read.csv("2_year_system_demand.csv")%>%
  rename('hourly_demand' = 'Hourly.Demand..cfs.',
         'date_time' = 'Read.Date.Time')%>%
  mutate(hourly_demand = as.numeric(hourly_demand),
         hourly_demand = ifelse(is.na(hourly_demand),0,hourly_demand)) #What do we do with NA's???

#Make Date and Time match a consistent format with the Weather Data
sd2 <- as.data.frame(str_split_fixed(sd$date_time, " ", 3))%>%
  rename(date = V1, time = V2, am_or_pm = V3)%>%
  mutate(datetime_str = paste(date, time, am_or_pm),
         datetime = as.POSIXct(strptime(datetime_str, format = "%m/%d/%y %I:%M %p")),
         datetime_military = strftime(datetime, format = "%m/%d/%Y %H:%M")) %>%
  select(datetime,datetime_military)
sd3 <-as.data.frame(str_split_fixed(sd2$datetime_military, " ", 2))%>%
  rename(date = V1, time = V2)
sd4 <-as.data.frame(str_split_fixed(sd3$date, "/", 3))%>%
  rename(CalendarMonth = V1,  CalendarDay = V2, CalendarYear = V3)
#Throw it all together
sd <- cbind(sd,sd2,sd3,sd4)%>%
  select(-date_time)
#Get rid of temp dfs
remove(sd2,sd3,sd4)

#The names and dates in this table do not need much conversion
wa <- read.csv("weather_sl_airport.csv")
  
  

sp <- read.csv("system_demand_by_point_type.csv")%>%
  rename('date_time' = 'Read.Date.Time',
         'well_flow_cfs' = 'Well.Flow..cfs.',
         'purchased_flow_cfs' = 'Purchased.Flow..cfs.',
         'transported_flow_cfs' = 'Transported.Flow..cfs.',
         'treated_flow_cfs' = 'Treated.Flow..cfs.',
         'injected_flow_cfs' = 'Injected.Flow..cfs.',
         'reservoir_trend_cfs' = 'Reservoir.Trend..cfs.',)
#Make Date and Time match a consistant format with the Weather Data
sp2 <- as.data.frame(str_split_fixed(sp$date_time, " ", 3))%>%
  rename(date = V1, time = V2, am_or_pm = V3)%>%
  mutate(datetime_str = paste(date, time, am_or_pm),
         datetime = as.POSIXct(strptime(datetime_str, format = "%m/%d/%y %I:%M %p")),
         datetime_military = strftime(datetime, format = "%m/%d/%Y %H:%M")) %>%
  select(datetime,datetime_military)
sp3 <-as.data.frame(str_split_fixed(sp2$datetime_military, " ", 2))%>%
  rename(date = V1, time = V2)
sp4 <-as.data.frame(str_split_fixed(sp3$date, "/", 3))%>%
  rename(CalendarMonth = V1,  CalendarDay = V2, CalendarYear = V3)
#Throw it all together
sp <- cbind(sp,sp2,sp3,sp4)%>%
  mutate(reservoir_trend_cfs = as.numeric(reservoir_trend_cfs))%>%
  select(-date_time)
#Get rid of temp dfs
remove(sp2,sp3,sp4)
```
> Something important to note is that the weather data is daily and demand data is hourly so we should ask if they want to change to hourly weather data or as how we want to agregate daily demand so we can match the grain for our model.

>Also note below some errors found and questions raised on initial data inspection

```{r}
#It appears that there are 3 instances with negative hourly demand.
#Does this actually mean that water was given back or is this an error?
sd %>% 
  filter(hourly_demand<0)

#Is there an entry for each hour of each day of each year?
sd %>% 
  group_by(CalendarDay) %>% 
  summarize(count=n())
#No there is not. We are missing 2 hours on the 8th and 14th of some year
sd[sd$time %in% "",]
#It looks like row 1610 show that we don't have data for 01h00 - 04h00 on 3/8/2020
#one row is NA, and there is a missing column
#And the exact same gap on row 10513 03/14/2021

#Is there possibly some planned maintenance going on in early March of each year?
```


## Add a season variable to all tables

```{r}
sd <- sd%>%
  mutate(season = ifelse(CalendarMonth %in% c('03','04','05'), 'Spring',
                         ifelse(CalendarMonth %in% c('06','07','08'), 'Summer',
                                ifelse(CalendarMonth %in% c('09','10','11'), 'Fall',
                                       ifelse(CalendarMonth %in% c('12','01','02'), 'Winter','0')))))%>%
  filter(season != '0')


sp <- sp%>%
  mutate(season = ifelse(CalendarMonth %in% c('03','04','05'), 'Spring',
                         ifelse(CalendarMonth %in% c('06','07','08'), 'Summer',
                                ifelse(CalendarMonth %in% c('09','10','11'), 'Fall',
                                       ifelse(CalendarMonth %in% c('12','01','02'), 'Winter','0')))))%>%
  filter(season != '0')


wa <- wa%>%
  mutate(season = ifelse(CalendarMonth %in% c('3','4','5'), 'Spring',
                         ifelse(CalendarMonth %in% c('6','7','8'), 'Summer',
                                ifelse(CalendarMonth %in% c('9','10','11'), 'Fall',
                                       ifelse(CalendarMonth %in% c('12','1','2'), 'Winter','0')))))%>%
  filter(season != '0')
```

## System Demand joined to Weather Data from Salt Lake Airport

```{r}
#Snapshot View
sd1 <- sd%>%
  group_by(date)%>%
  summarise(daily_demand = mean(hourly_demand),
            CalendarMonth = CalendarMonth,
            CalendarDay = CalendarDay,
            CalendarYear = CalendarYear,
            season = season)

wa1 <- wa%>%
  select(-CalendarYear,-CalendarMonth,-CalendarDay,-season)%>%
  rename(date = WeatherDate)

wd <- merge(wa1,sd1,by="date")

# Factor RainDays
wd <- wd%>%
  mutate(RainDays = factor(RainDays))

#Get rid of temp dfs
remove(sd1,sd2,wa1)

summary(wd)

```

## Merge Data

```{r}
#Weather data begins prior to 2020 where the system data begins. Truncate for merging
merweth=wa %>% 
  filter(CalendarYear>=2020)
#Date column is a character. Convert to date
merweth=merweth %>% 
  mutate(WeatherDate = as.Date(WeatherDate,format="%m/%d/%Y"))
#Date column is in POSIXct format. Convert to date
mersd=sd %>% 
  mutate(ActualDate = as.Date(paste(as.numeric(sd$CalendarMonth),
                                    as.numeric(sd$CalendarDay),
                                    as.numeric(sd$CalendarYear),
                                    sep="-"),
                                    format="%m-%d-%Y"))
#Change hourly_demand to avg_daily demand for merge
mershrink=mersd %>% 
  group_by(ActualDate) %>% 
  summarize(avgdaily_demand = mean(hourly_demand))

#remove NA at the bottom
mershrink=mershrink[1:731,] 

#Note from Zach: We should change this so that it doesn't go by position so it can handle new data. Also should we replace the NA with mean or some other imputation or just get rid of it?

#Merge Datasets
merginated = merge(mershrink,merweth,by.x = "ActualDate",
                   by.y = "WeatherDate",all.y = TRUE,
                   all.x = FALSE)
```

## Add Hour and Min columns

```{r}
sd2 <- as.data.frame(str_split_fixed(sd$time,':',2))%>%
  rename(Hour = V1, Min = V2)

sd <- cbind(sd,sd2)

remove(sd2)
```


# 2 year system demand

```{r}
glimpse(sd)
summary(sd)
```

## Hourly Demand Histogram

```{r}
ggplot(sd,aes(hourly_demand))+
  geom_histogram()+
  theme_minimal()
```
> Demand seems to be heavily skewed towards the lower numbers. I wonder what happens if you change it to months or seasons.

## Hourly Demand Histogram by Calendar Month

```{r}
ggplot(sd,aes(hourly_demand))+
  geom_histogram()+
  theme_minimal()+
  facet_wrap(~CalendarMonth)
```


## Hourly Demand Histogram by season

```{r}
ggplot(sd,aes(hourly_demand))+
  geom_histogram()+
  theme_minimal()+
  facet_wrap(~season)

#What are the x axis values in each of these histograms?

```

## Hourly Demand Histogram by Calendar year

```{r}
ggplot(sd,aes(hourly_demand))+
  geom_histogram()+
  theme_minimal()+
  facet_wrap(~CalendarYear)
```

> The difference between 2020 and 2021 don't seem very notable. 
> This is super interesting considering 2021 was considered a major drought year. Does look like the demand in 2021 was a bit higher surprisingly.



## Average Hourly Demand

```{r}
sd%>%
  group_by(Hour)%>%
  summarise(mean(hourly_demand))%>%
  rename(avg_demand = 'mean(hourly_demand)')%>%
  ggplot(aes(Hour, avg_demand))+
  geom_col()+
  theme_minimal()
```
> Demand seems to drop towards the middle of the day. This makes sense as people typically water during the early morning hours

## Average Hourly Demand by Season

```{r}
sd%>%
  group_by(season,Hour)%>%
  summarise(mean(hourly_demand))%>%
  rename(avg_demand = 'mean(hourly_demand)')%>%
  ggplot(aes(Hour, avg_demand))+
  geom_col()+
  theme_minimal()+
  facet_wrap(~season)
```

> Seems like the Demand is higher in the summer as expected and seems to be higher earlier in the day.

## Average Season Demand

```{r}
sd%>%
  group_by(season)%>%
  summarise(mean(hourly_demand))%>%
  rename(avg_demand = 'mean(hourly_demand)')%>%
  ggplot(aes(season, avg_demand))+
  geom_col()+
  theme_minimal()
```
> The average demand is way higherin the summer and the next highest is fall.

## Average Monthly Demand

```{r}
sd%>%
  group_by(CalendarMonth)%>%
  summarise(mean(hourly_demand))%>%
  rename(avg_demand = 'mean(hourly_demand)')%>%
  ggplot(aes(CalendarMonth, avg_demand))+
  geom_col()+
  theme_minimal()
```
## Average Hourly Demand by Year

```{r}
sd%>%
  group_by(CalendarYear)%>%
  summarise(sum(hourly_demand))%>%
  rename(Sum_of_demand = 'sum(hourly_demand)')%>%
  ggplot(aes(CalendarYear, Sum_of_demand))+
  geom_col()+
  theme_minimal()
```

> It looks like 2020 had slightly more water demand.


# System Demand By Point Type

```{r}
glimpse(sp)
summary(sp)
```
## Average Well Flow by Season

```{r}
sp%>%
  group_by(season)%>%
  summarise(mean(well_flow_cfs))%>%
  rename(avg_demand = 'mean(well_flow_cfs)')%>%
  ggplot(aes(season, avg_demand))+
  geom_col()+
  theme_minimal()
```

## Average Purchased Flow by Season

```{r}
sp%>%
  group_by(season)%>%
  summarise(mean(purchased_flow_cfs))%>%
  rename(avg_demand = 'mean(purchased_flow_cfs)')%>%
  ggplot(aes(season, avg_demand))+
  geom_col()+
  theme_minimal()
```
## Average Transported Flow by Season

```{r}
sp%>%
  group_by(season)%>%
  summarise(mean(transported_flow_cfs))%>%
  rename(avg_demand = 'mean(transported_flow_cfs)')%>%
  ggplot(aes(season, avg_demand))+
  geom_col()+
  theme_minimal()
```
## Average Treated Flow by Season

```{r}
sp%>%
  group_by(season)%>%
  summarise(mean(treated_flow_cfs))%>%
  rename(avg_demand = 'mean(treated_flow_cfs)')%>%
  ggplot(aes(season, avg_demand))+
  geom_col()+
  theme_minimal()
```

## Average Injected Flow by Season

```{r}
sp%>%
  group_by(season)%>%
  summarise(mean(injected_flow_cfs))%>%
  rename(avg_demand = 'mean(injected_flow_cfs)')%>%
  ggplot(aes(season, avg_demand))+
  geom_col()+
  theme_minimal()
```

> Injected flow seems to only happen durring the winter

## Injected Flow Histogram

```{r}
ggplot(sp,aes(injected_flow_cfs))+
  geom_histogram()+
  theme_minimal()
```

> Seems like it's usually 0.

## Average Reservoir Flow by Season

```{r}
sp%>%
  group_by(season)%>%
  summarise(sum(reservoir_trend_cfs))%>%
  rename(sum_of_demand = 'sum(reservoir_trend_cfs)')%>%
  ggplot(aes(season, sum_of_demand))+
  geom_col()+
  theme_minimal()
```

> I did sum on this one becase it had a lot of weird negative and positive and I wanted to see how it corelated

## Reservoir Flow Histogram

```{r}
ggplot(sp,aes(reservoir_trend_cfs))+
  geom_histogram()+
  theme_minimal()
```


# Weather Data from Salt Lake Airport

```{r}
glimpse(wa)
summary(wa)
```


## Check for useful predictors
```{r}
#Clear out redundant columns
merginated=merginated[,c(1,2,8,9,12,13,14)]
mermod<-lm(avgdaily_demand~.,data = merginated)
summary(mermod)
#Here we can see that Precipitation has a p value of 0.09. logically rain
#coming same day should not affect as much as recent days (within a week or so)

#Create ISO week number identifier column
merginated = merginated %>% 
  mutate(weeknum = ISOweek(ActualDate))
#Create list to show precipitation for a given ISO week number
weekref = merginated %>% 
  group_by(weeknum) %>% 
  summarize(weekfall = sum(Precipitation))
#Visualize
weekref %>% 
  ggplot(aes(weeknum,weekfall))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle("Precipitation by Week Number")
#Seems a bit sporadic

#shift column so that list shows prior week precipitation
weekref = shift.column(weekref,columns = "weekfall",len = 1L,up=TRUE)
weekref = weekref[,c(1,3)]
weekref = weekref %>% 
  rename(prevweekfall = weekfall.Shifted)
#Merge list to create column showing rainfall by previous week
Merged_Data_Prevweek = merge(merginated,weekref,by.x = "weeknum",by.y = "weeknum",all.y = TRUE,all.x = TRUE)
mergiweek=Merged_Data_Prevweek[,c(2,3,4,6,7,8,9)]
merwod<-lm(avgdaily_demand~.,data = mergiweek)
#Simple regression showing variables with previous week rainfall included
summary(merwod)
#Does seem to be significant

#Remove temp data frames
remove(mergiweek,weekref,merwod,mermod,mersd,merweth,mershrink,merginated)
```
>Perhaps we should explore using previous week's precipitation data for prediction.
Or in some other way reference recent precipitation history instead of same day history

## Average Demand by TemperatureMean

```{r}
wd%>%
  group_by(TemperatureMean)%>%
  summarise(mean(daily_demand))%>%
  rename(avg_demand = 'mean(daily_demand)')%>%
  ggplot(aes(TemperatureMean, avg_demand))+
  geom_line()+
  theme_minimal()
```


## Average Demand by Depth of Snow

```{r}
wd%>%
  group_by(SnowDepth)%>%
  summarise(mean(daily_demand))%>%
  rename(avg_demand = 'mean(daily_demand)')%>%
  ggplot(aes(SnowDepth, avg_demand))+
  geom_col()+
  theme_minimal()
```

## Average Demand on rainy days vs non rainy

```{r}
wd%>%
  group_by(RainDays)%>%
  mutate(count_of_days = 1)%>%
  summarise(count_of_days = sum(count_of_days),
            mean(daily_demand))%>%
  rename(avg_demand = 'mean(daily_demand)')
```

## Average Demand on Snow days vs non Snow days

```{r}
wd%>%
  mutate(snow_day = ifelse(SnowFall>0,1,0))%>%
  group_by(snow_day)%>%
  mutate(count_of_days = 1)%>%
  summarise(count_of_days = sum(count_of_days),
            mean(daily_demand))%>%
  rename(avg_demand = 'mean(daily_demand)')
```

# Numeric Variable Examination 

## Relation to Daily Demand

```{r}
wdn <- wd%>%
  select_if(is.numeric)

#an attempt to manually look at the most important numberic vars
cwdn <- cor(wdn, use="pairwise.complete.obs")
cwdn1 <- data.frame(sort(cwdn[,'daily_demand'], decreasing = TRUE))

cwdn1%>%
  rename(Correlation = "sort.cwdn....daily_demand....decreasing...TRUE.")
```

```{r}


#Correlation Plot

#Calculate correlation
dfcor <- cor(wdn[,unlist(lapply(df, is.numeric))])

#Plot the correlation

#Only Select Numeric Variables
df_numericVars <- which(sapply(wdn, is.numeric)) 


#correlations of all numeric variables
df_numVar <- wdn[, df_numericVars]
df_cor_numVar <- cor(df_numVar, use="pairwise.complete.obs") 



# Correlation of numeric variables with their numbers and shap/color relation to sale price
corrplot.mixed(df_cor_numVar, tl.col="black", tl.pos = "lt", tl.cex = 0.7,cl.cex = .7, number.cex=.7)

corrplot(df_cor_numVar, tl.col="black", tl.pos = "lt", tl.cex = 0.7,cl.cex = .7, number.cex=.7, method = "number")
```










## Correlation of logged mean temprature with logged daily demand

```{r}
ggplot(wdn, aes(log(TemperatureMean),log(daily_demand)))+
  geom_point()+ stat_smooth(method="lm", se=FALSE)

```

## Correlation of logged Precipitation with logged daily demand

```{r}
ggplot(wdn, aes(log(Precipitation),log(daily_demand)))+
  geom_point()+ stat_smooth(method="lm", se=FALSE)

```

## Correlation of logged SnowFall with logged daily demand

```{r}
ggplot(wdn, aes(log(SnowFall),log(daily_demand)))+
  geom_point()+ stat_smooth(method="lm", se=FALSE)

```



## Correlation of logged CoolingDegreeDays with logged daily demand

```{r}
ggplot(wdn, aes(log(CoolingDegreeDays),log(daily_demand)))+
  geom_point()+ stat_smooth(method="lm", se=FALSE)

```


# Results and Conclusions

During this EDA we discovered that are initial assumption holds true. Temperature plays an important role in water usage. There is more demand on hotter days and seasons when there isn't precipitation. We can see that there are correlations between daily demand and Temperature Mean, Min, and Max, CoolingDegreeDays, SnowDepth, Precipitation, and Snowfall. The weather data available currently only tracks metrics on a daily basis. The data received from JVWCD tracks demand minute by minute.We then merged and aggregated the data to fit our needs. We ran a linear regression where we discovered that the previous week precipitation is significant along with temperature and therefore seasons.Additionally, we see how the different point types are affected by the weather data as well and there overall distribution. Our initial goal is to successfully predict daily demand and then see how that compares with the necessary point types. 





# Questions for Jordan Valley
Difference in weather and water granularity?
What do we do with NULLs?
What do we do with negative hourly demand values?
Missing hourly entries?
If a lot of this data makes logical sense (meaning we can generally predict demand based on past data and an upcoming weather forecast), what additional value is Jordan Valley looking for in a prediction model?


# Member Contributions

Jake: 
Added correlations plots
Formatted final report
Added business problem, and analytical objective

Nate:
Included highlights of N/A and negative values to discuss
Included merged data frame showing previous week's precipitation to check for prediction power of recent rainfall

Zach:

Jackson: